---
title: "Package arlab4"
author: "Josh Hyungyum Kim"
date: "2017 09 24"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  library(gridExtra)
)
```
##Introduction
 The main purpose of this vignette is to describe and elaborate the package arlab4 which contains class  linreg and its  methods. Basically, linreg is the Reference Class and it has 6 methods regarding multiple linear regression. It takes 2 arguments, formula and data. The formula is a formula for your regression process and the data should be a data.frame for the regression.
 
##Methods
 6 methods are included in this linreg class. 
 
###print()
Print out the coefficients and coefficient names.

###plot()
Plotting Residuals vs Fitted graph & Scale-Location graph.

###resid()
Return the vector of residuals e.

###pred()
Return the predicted values y_hat.

###coef()
Return the coefficients as a named vector

###summary()
Simplified summary of the linear model.

###Example output of method print()
```{r, echo=FALSE, eval=TRUE}
linreg <- setRefClass("linreg", 
  fields = list(formula="formula",
                data="data.frame",
                reg_coe="matrix",
                fit_val="matrix",
                residu="matrix",
                dof="numeric",
                res_var="matrix",
                var_reg_coe="numeric",
                t_val="matrix",
                data_name="character"),
  
  methods = list(
    initialize = function(formula, data){
      
      formula <<- formula
      data <<- data
      
      data_name <<- deparse(substitute(data))
      
      #getting y variable from the formula
      get_y <- all.vars(formula)[1]
      
      #defining X and y
      y <- data[[get_y]]
      X <- model.matrix(formula, data)
      
      reg_coe <<- solve((t(X)%*%X))%*%t(X)%*%y
      fit_val <<- X%*%reg_coe
      residu <<- y-fit_val
      
      dof <<- length(y) - length(all.vars(formula))
      res_var <<- (t(residu)%*%residu)/dof
      var_reg_coe <<- diag(as.numeric(res_var)*solve(t(X)%*%X))
      t_val <<- reg_coe/sqrt(var_reg_coe)
       
    },
  
    print = function(){
      "Print out the coefficients and coefficient names"
      
      cat("Call: \n ")
      cat(paste0("linreg(formula = ",format(formula),", data = ",data_name,")\n\n"))
      cat("Coefficients: \n")
      cat(" ",row.names(reg_coe), "\n    ", sep = "  ")
      cat(t(reg_coe), sep="    ")
     
    }
  )
  )
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length, data=iris)
linreg_mod$print()
```

###Example plot of method plot()
```{r, echo=FALSE, eval=TRUE, fig.width=7}
linreg <- setRefClass("linreg", 
  fields = list(formula="formula",
                data="data.frame",
                reg_coe="matrix",
                fit_val="matrix",
                residu="matrix",
                dof="numeric",
                res_var="matrix",
                var_reg_coe="numeric",
                t_val="matrix",
                data_name="character"),
  
  methods = list(
    initialize = function(formula, data){
      
      formula <<- formula
      data <<- data
      
      data_name <<- deparse(substitute(data))
      
      #getting y variable from the formula
      get_y <- all.vars(formula)[1]
      
      #defining X and y
      y <- data[[get_y]]
      X <- model.matrix(formula, data)
      
      reg_coe <<- solve((t(X)%*%X))%*%t(X)%*%y
      fit_val <<- X%*%reg_coe
      residu <<- y-fit_val
      
      dof <<- length(y) - length(all.vars(formula))
      res_var <<- (t(residu)%*%residu)/dof
      var_reg_coe <<- diag(as.numeric(res_var)*solve(t(X)%*%X))
      t_val <<- reg_coe/sqrt(var_reg_coe)
       
    },
  
    plot = function(){
      "Plotting Residuals vs Fitted graph & Scale-Location graph"
      
      library(ggplot2)
      library(gridExtra)
      
      liu_theme <- 
        theme(
          #margin setting
          plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
          
          #color codes from: https://html-color.codes/hex
          
          #background color Liu-base color 9
          panel.background = element_rect(fill="#a5f0e4"),
          
          #delete grid lines
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          
          #text color of title and axis Liu-accent color 1
          plot.title = element_text(color="#ff6442", face = "bold", size="12"),
          axis.title.x = element_text(color="#ff6442", size="12"),
          axis.title.y = element_text(color="#ff6442", size="12"),
          axis.text.x = element_text(color="#ff6442", size="9"),
          axis.text.y = element_text(color="#ff6442", size="9"),
          
          #border setting
          panel.border = element_rect(fill = NA),
          
          #caption color Liu_base color 2
          plot.caption = element_text(color="#17c7d2", face="bold", size="9")
        ) 
      
      #plot1
      plot1 <- ggplot(data.frame(fit_val,residu),aes(y=residu,x=fit_val)) + geom_point() + #graph + scatter plot
        xlab(paste("Fitted values\n", "lm(",format(formula),")")) + 
        ylab("Residuals")+ ggtitle("Residuals vs Fitted") + 
        geom_smooth(span = 2,colour="red",method="loess",se=FALSE) + labs(caption="2017 LiU") +
        theme(plot.title = element_text(hjust = 0.5)) + #Title center alignment
        liu_theme #Theme added
        
      #plot2
      std_res <- residu/sd(residu)
      plot2 <- ggplot(data.frame(fit_val,std_res),aes(y=std_res,x=fit_val)) + geom_point() + #graph + scatter plot
        xlab(paste("Fitted values\n", "lm(",format(formula),")")) +
        ylab(expression(sqrt(abs("Standardized residuals")))) + ggtitle("Scale-Location") +
        geom_smooth(span = 2,colour="red",method="loess",se=FALSE) + labs(caption="2017 LiU") +
        theme(plot.title = element_text(hjust = 0.5)) + #Title center alignment
        liu_theme #Theme added
        
      grid.arrange(plot1, plot2, ncol=2)
    }
  )
  )
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length, data=iris)
linreg_mod$plot()
```

###Example output of method summary()
```{r, echo=FALSE, eval=TRUE}
linreg <- setRefClass("linreg", 
  fields = list(formula="formula",
                data="data.frame",
                reg_coe="matrix",
                fit_val="matrix",
                residu="matrix",
                dof="numeric",
                res_var="matrix",
                var_reg_coe="numeric",
                t_val="matrix",
                data_name="character"),
  
  methods = list(
    initialize = function(formula, data){
      
      formula <<- formula
      data <<- data
      
      data_name <<- deparse(substitute(data))
      
      #getting y variable from the formula
      get_y <- all.vars(formula)[1]
      
      #defining X and y
      y <- data[[get_y]]
      X <- model.matrix(formula, data)
      
      reg_coe <<- solve((t(X)%*%X))%*%t(X)%*%y
      fit_val <<- X%*%reg_coe
      residu <<- y-fit_val
      
      dof <<- length(y) - length(all.vars(formula))
      res_var <<- (t(residu)%*%residu)/dof
      var_reg_coe <<- diag(as.numeric(res_var)*solve(t(X)%*%X))
      t_val <<- reg_coe/sqrt(var_reg_coe)
       
    },
  
    summary = function(){
      "Simplified summary of the linear model"
      
      #p_value calculation
      p_val <- 2*pt(abs(t_val),dof,lower.tail = FALSE)
      
      #organizing table
      reg_coe_edit <- matrix(NA, nrow=3, ncol=4)
      reg_coe <<- cbind(reg_coe, reg_coe_edit)
      reg_coe[,2] <<- sqrt(var_reg_coe)
      reg_coe[,3] <<- t_val
      reg_coe[,4] <<- p_val
      reg_coe[,5] <<- "***"

      cat("          ", c("Estimate", "Std. Error", "t value", "Pr(>|t|)", ""), "\n", sep = "        ")
      for(i in 1:length(all.vars(formula))){
      cat(row.names(reg_coe)[i], reg_coe[i,], "\n", sep = " ")
      }

      cat("\nResidual standard error:",format(sd(residu)), "on 147 degrees of freedom\n\n")
      }
  )
  )
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length, data=iris)
linreg_mod$summary()
```